---
title: "eds221_day4_comp"
author: "Kiran Favre"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Lecture examples

### for loop examples

```{r}
dosage <- c(10, 100, 1000)

dose_index <- vector(mode = "character", length = length(dosage))

for (i in seq_along(dosage)) {
  k <- paste0("zinc_", dosage[i])
  dose_index[i] <- k
}

dose_index[i]
```

### A nested for loop example

```{r}
file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

for(i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}
```

## Functions

```{r}
exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}

exclaim_age(age=22)
```

## Interactive session

### Warm up

```{r}
quarter_splits <- c( 1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

## Run a for loop that returns your total time for any two consecutive quarter miles. hint: the first is 1.0+1.1, then 1.1+1.2, etc

for (i in seq_along(quarter_splits)) {
  half_time <- quarter_splits[i] + quarter_splits[i + 1]
  print(half_time)
}
```

### Function

```{r}
animal_age <- function(animal, age) {
  if (animal == "dog") {
    print(age * 7)
  } else if (animal == "goat") {
    print(age * 4.7)
  }
}
animal_age("dog", 7)
```

Make a function to double the user input

```{r}
double_this <- function(my_value) {
  print(2* my_value)
}

double_this(my_value = 200)
double_this(200)

double_this(c(1, 2, 10, 100, 175))
```

Write a function that asks for a user's name and age and returns an exclamation "My name is [name], and I am [age] years old."

```{r}
exclaim_name_age<- function(your_name, your_age) {
  print(paste("My name is", your_name, "and I am", your_age, "years old!"))
}

exclaim_name_age("Kiran", 22)
```

### Conditionals within functions

```{r}
convert_animal_age <- function(animal, human_yrs) {
  if (animal == "dog") {
    print(human_yrs * 7)
  } else if (animal == "goat") {
    print(human_yrs * 4.7)
  } else {
    print("Please enter dog or goat for animal.")
    }
}

convert_animal_age("dog", 12)
convert_animal_age("goat", 22)
convert_animal_age("puppy", 21)
```

Data frame input

```{r}
dog_choice <- tribble(
  ~ dog_name, ~ food, 
  "Batman", "yogurt",
  "Robin", "cheese",
  "Ozzie", "almonds",
  "Brandi", "pizza"
)

#filtering dog choice to only keep observations that keeps dog name values that match what the user chooses
dog_menu <- function(name) {
  my_sub <- filter(dog_choice,
                   dog_name == name)
  print(paste("My name is",
              my_sub$dog_name,
              "and my favorite food is",
              my_sub$food))
}

dog_menu(name= "Batman")
dog_menu(name = "Brandi")
dog_menu(name = "Robin")
```

### Functions meet for loops

```{r}

## for the sequenece along this length, we will find the means of the values in the columns, define the name so we can use this in our output, and then print those in a statement
## colnames is interchangeable with names

means_over_columns <- function(dataframe){
  for(i in seq_along(dataframe)) {
    col_mean <- mean(dataframe[[i]],
                     na.rm = TRUE)
    column_names <- colnames(dataframe[i])
    print(paste("the mean value of", column_names, "is", col_mean))
  }
}

means_over_columns(dataframe = mtcars)

means_over_columns(dataframe = faithful)
```

# Afternoon session

comprehension check:

this will produce a 3x4 matrix

this looks at i=1, j=1-4. then i=2, j=1-4. then i=3, j=1-4. then i=4, j=1-4

```{r}
apples <- c(1, 2, 3, 4)
price <- c(7, 10, 25)

fruit_out <- matrix(nrow = length(price), ncol = length(apples))

for (i in seq_along(apples)) {
  for (j in seq_along(price)) {
    total_cost <- price[j] *apples[i]
    fruit_out[j,i] <- total_cost
  }
}
```

Our own examples

```{r}
clothing <- c("shirt", "pants", "shoes")
color <- c("purple", "red", "blue", "white")


our_clothes <- matrix(nrow = length(color), ncol = length(clothing))
# clothing will be populated by column var and color by row

for (i in seq_along(clothing)) {
  for (j in seq_along(color)) {
    clothing_details <- paste(color[j], clothing[i])
    our_clothes[j, i] <- clothing_details
  }
}

our_clothes
```

```{r}
first_name <- c("kiran", "elise", "Michael")
last_name <- c("chair", "table", "light")

full_names <- matrix(nrow = length(first_name), ncol = length(last_name))

for (i in seq_along(first_name)) {
  for (j in seq_along(last_name)) {
    add_names <- paste(first_name[i], last_name[j])
    full_names[j,i] <- add_names
  }
}

full_names
```
